A Simplified Plackett-Luce Likelihood Function for Rank Orders

### 2025-04-01 inst/hankin-AustrianJournalofStatistics.tex
issue #239: easily-fixed referee comments

\begin{eqnarray}\label{likeforrn1}
  \mathcal{L}_{r,n}(a) &=&
\frac{b}{1+ (n-2)b}\cdot
\frac{b}{a+(n-3)b}\cdots\frac{a}{a+(n-r-1)b}\nonumber\\
&=& \frac{B-1}{(B+n-1)(B+n-2)\cdots(B+n-r-1)}\nonumber\\ 
&=& \frac{(B-1)(B+n-r-2)!}{(B+n-1)!}\label{likeforrn3}
\end{eqnarray}

@@

\begin{table}[t]
  \centering
  \caption{Parkrun results}
\label{parkruntable}
\begin{tabular}{cccccccccc}\\
 & \multicolumn{9}{c}{2023}\\
rank   & 177& 222& 206& 142& 118& 224& 128& 115& 183\\
runners& 259& 305& 297& 241& 179& 338& 203& 245& 254\\ \\
\end{tabular}
\begin{tabular}{ccccccccccccc}
  & \multicolumn{12}{c}{2024}\\
rank   &  193& 240& 197& 257& 227& 234& 238& 238& 196& 242& 242& 318\\
runners&  145& 157& 156& 191& 172& 174& 173& 165& 172& 199& 181& 229\\
\end{tabular}
\end{table}

@@

One might wonder whether the student was better at pure or applied
mathematics and, if so, to quantify the difference.
Figure~\ref{plotpureandapplied} shows contours of support [again
  normalised so the support of the evaluate is zero] for the dataset,
as a function of pure strength on the horizontal axis and applied
strength on the vertical axis.  A null of equal strengths appears as
the diagonal line.  We see that the maximum likelihood estimate is
$\hat{p}_\mathrm{pure}=0.542$, $\hat{p}_\mathrm{applied}=0.852$.
However, a constrained optimization shows that the maximum support
along the null diagonal is about $-2.06$, at
${p}_\mathrm{pure}={p}_\mathrm{applied}\simeq 0.683$.  The two units of
support criterion~\citep{edwards1992} is met, and we are justified in
concluding that the student is indeed more competent at applied than
pure mathematics.  Alternatively, we may observe that
$2\mathcal{S}=4.11$ is in the tail region of its asymptotic $\chi^2_1$
distribution, corresponding to a $p$-value of about $0.042$.

>>>>

\begin{eqnarray}\label{likeforrn1}
  \mathcal{L}_{r,n}(a) &=&
\frac{b}{1+ (n-2)b}\cdot
\frac{b}{a+(n-3)b}\cdots\frac{a}{a+(n-r-1)b}\nonumber\\
&=& \frac{B-1}{(B+n-2)\cdots(B+n-r-1)}\nonumber\\ 
&=& \frac{(B-1)(B+n-r-2)!}{(B+n-1)!}\label{likeforrn3}
\end{eqnarray}

@@

\begin{table}[t]
  \centering
  \caption{Parkrun results}
\label{parkruntable}
\begin{tabular}{cccccccccc}\\
 & \multicolumn{9}{c}{2023}\\
rank   & 177& 222& 206& 142& 118& 224& 128& 115& 183\\
runners& 259& 305& 297& 241& 179& 338& 203& 245& 254\\ \\
\end{tabular}
\begin{tabular}{ccccccccccccc}
  & \multicolumn{12}{c}{2024}\\
rank      &  145& 157& 156& 191& 172& 174& 173& 165& 172& 199& 181& 229\\
runners   &  193& 240& 197& 257& 227& 234& 238& 238& 196& 242& 242& 318\\
\end{tabular}
\end{table}

@@

One might wonder whether the student was better at pure or applied
mathematics and, if so, to quantify the difference.
Figure~\ref{plotpureandapplied} shows contours of support [again
  normalised so the support of the evaluate is zero] for the dataset,
as a function of pure strength on the horizontal axis and applied
strength on the vertical axis.  A null of equal strengths appears as
the diagonal line.  We see that the maximum likelihood estimate is
$\hat{p}_\mathrm{pure}=0.542$, $\hat{p}_\mathrm{applied}=0.852$.
However, a constrained optimization shows that the maximum support
along the null diagonal is about $-2.06$, at
${p}_\mathrm{pure}={p}_\mathrm{applied}\simeq 0.683$.  The two units
of support criterion~\citep{edwards1992} is met, and we are justified
in concluding that the student indeed has a higher ranking in the
applied modules than the pure modules.  Alternatively, we may observe
that $2\mathcal{S}=4.11$ is in the tail region of its asymptotic
$\chi^2_1$ distribution, corresponding to a $p$-value of about
$0.042$.
